Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-16 paper/us_election_2020.qmd
some edits

The remainder of this paper is structured into different sections. @sec-data demonstrates the data used for our report and includes some tables and graphs to illustrate the different groups of people in our data. @sec-model builds the model and discusses its justification and explanation. @sec-result highlights the results of the predictions using tables and graphs. @sec-discussion contains discussions that conducted based on the findings, which addresses the voting prediction results based on race, region, employment status, and the influence of COVID-19 and in-mail voting systems.

@@

1. votereg: whether the respondent is registered to vote;

2. CC20_410: the presidential candidate the respondent voted for;

3. region: the census region where the respondent lives; options are "Northeast", "Midwest", "South", and "West".

4. race: the racial or ethnic group of the respondent; options are "White", "Black", "Hispanic", "Asian", "Native American", "Middle Eastern", "Two or more races", and "Other".

5. employ: the current employment status of the respondent; options are "Full-time", "Part-time", "Temporarily laid off", "Unemployed", "Retired", "Permanently disabled", "Homemaker", "Student", and "Other".

@@

The dataset is cleaned by renaming the column names, specifying the class of the columns, and changing the numbers in the table to the corresponding description in the codebook to improve the readability. In addition, NA employment status is removed in the cleaned data file since there are only 34 entries, which is a small amount of missing value compared with 61000 entries. After cleaning, 43534 rows of data with 4 data features remain.

The cleaned data will be analyzed and performed using R [@citeR] with `tidyverse` [@citeTidyverse], `here` [@citeHere], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModleSummary], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], `marginaleffects` [@citeMarginalEffects], `plotly` [@citePlotly], `tibble` [@citeTibble], `margins` [@citeMargins], `testthat` [@citetestthat] and `kableExtra` [@citeKableExtra]. @tbl-clean-data shows a preview of the cleaned dataset.

>>>>

The remainder of this paper is structured into different sections. @sec-data demonstrates the data used for our report and includes some tables and graphs to illustrate the different groups of people in our data. @sec-model builds the model and discusses its justification and explanation. @sec-result highlights the results of the predictions using tables and graphs. @sec-discussion contains discussions that conducted based on the findings, which addresses the voting prediction results based on race, region, employment status, and the influence of COVID-19 and in-mail voting systems. Statistical programming language R [@citeR] is used in this report, with packages `tidyverse` [@citeTidyverse], `here` [@citeHere], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModleSummary], `ggplot2` [@citeGgplot2], `knitr` [@citeKnitr], `marginaleffects` [@citeMarginalEffects], `plotly` [@citePlotly], `tibble` [@citeTibble], `margins` [@citeMargins], `testthat` [@citetestthat] and `kableExtra` [@citeKableExtra].

@@

1. votereg: whether the respondent is registered to vote; numeric numbers starting from 1 correspond to options "Yes", "No", and "Don't know", respectively.

2. CC20_410: the presidential candidate the respondent voted for; numeric numbers starting from 1 correspond to  options "Joe Biden (Democrat)", "Donald J. Trump (Republican)", "Other", "I did not vote in this race", "I did not vote", "Not sure", respectively.

3. region: the census region where the respondent lives; numeric numbers starting from 1 correspond to options "Northeast", "Midwest", "South", and "West", respectively.

4. race: the racial or ethnic group of the respondent; numeric numbers starting from 1 correspond to options "White", "Black", "Hispanic", "Asian", "Native American", "Middle Eastern", "Two or more races", and "Other", respectively.

5. employ: the current employment status of the respondent; numeric numbers starting from 1 correspond to options "Full-time", "Part-time", "Temporarily laid off", "Unemployed", "Retired", "Permanently disabled", "Homemaker", "Student", and "Other", respectively.

@@

The dataset is cleaned by renaming the column names, specifying the class of the columns, and changing the numbers in the table to the corresponding description in the codebook to improve the readability. In addition, NA employment status is removed in the cleaned data file since there are only 34 entries, which is a small amount of missing value compared with 61000 entries. After cleaning, 43534 rows of data with 4 data features remain. @tbl-clean-data shows a preview of the cleaned dataset.
